DATA 621 - Business Analytics and Data Mining

### 2020-11-08 HW4/HW4_Draft.Rmd
Minor cleanup, adding comments, calling arm library, slecting corrplot's corrplot

```{r loadData, include=FALSE}
# Load necessary libraries
library(corrplot)
library(ggplot2)
library(scales)
library(knitr)
library(caret)
library(RANN)
library(pROC)
library(data.table)
library(mice)

@@

```{r corrgram, fig.width=6, fig.height=4}
corrplot(cor(DT[, ..nmtrnNUM], use = 'complete.obs'),
         method = 'ellipse', type = 'lower', order = 'hclust',
         hclust.method = 'ward.D2')
```

@@

```{r trainTestSplit}
set.seed(642)
trnIDX <- createDataPartition(DT$TARGET_FLAG, p = 0.7)
trnX <- DT[trnIDX$Resample1, ]
tstX <- DT[!trnIDX$Resample1, ]
```

@@

```{r model1addVars}
# Add OWN variable for model set 1
trnX[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0),
                     levels = c(0, 1), labels = c('No', 'Yes'))]
tstX[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0),
                     levels = c(0, 1), labels = c('No', 'Yes'))]
```

## Model Set #2  
### Missing Data

Like Model Set #1, the data preparation efforts for Model Set #2 also includes the 
imputation of missing values for the predictor variables. However, because it combines 
the standard linear regression and the nearest-neighbor imputation approach, Model Set #2 
elected to fill in missing values of a continuous variable by using the predictive mean 
matching imputation method. Moreover, there will be the derivation of three new binary 
factor variables.

@@

### New Variables

For Model Set #2, new factor variables ($VARIABLE`.f`) were create as a way to better 
capture policyholder's profile in the binary logistic regression modeling process.

@@

```{r m2clean}
# Clean Training set
m2.train.X <- trnX
m2.train.X$HOME_VAL[is.na(m2.train.X$HOME_VAL)] <- 0
m2.train.X$OWN[is.na(m2.train.X$OWN)] <- "No"
m2.train.X$CAR_AGE = abs(m2.train.X$CAR_AGE) 
m2.train.r <- factor(m2.train.X$TARGET_FLAG, levels = c(0, 1), 
                     labels = c("NoClaim", "Claim"))
m2.train.X.impute <- mice(m2.train.X, method = 'pmm', print = FALSE)
m2.train.X <- complete(m2.train.X.impute)
m2.train.X$CAR_AGE.f <- factor(ifelse(m2.train.X$CAR_AGE > 1, 1, 0),
                              levels = c(0, 1), 
                              labels = c('1 Year or less', 'More than a Year'))
m2.train.X$HOMEKIDS.f <- factor(ifelse(m2.train.X$HOMEKIDS >= 1, 1, 0),
                              levels = c(0, 1), 
                              labels = c('None', 'At least 1'))
m2.train.X$KIDSDRIV.f <- factor(ifelse(m2.train.X$KIDSDRIV >= 1, 1, 0),
                              levels = c(0, 1), 
                              labels = c('None', 'At least 1'))
m2.train.p <- m2.train.X[, -c(1:3)]

@@

```{r model1DummyFreq}
# Cast target as factor for caret purposes
trnFreq <- factor(trnX$TARGET_FLAG, levels = c(0, 1),
                  labels = c("NoClaim", "Claim"))
# Set up dummy variables
trnDumFreq <- dummyVars(TARGET_FLAG ~ . +
                          OWN * HOME_VAL * INCOME +
                          CLM_FREQ * OLDCLAIM * MVR_PTS +
                          CAR_USE * URBANICITY -
                         INDEX - TARGET_AMT,
                        fullRank = TRUE, data = trnX)
# Create matrix of dummy values
trnXFD <- predict(trnDumFreq, newdata = trnX)
```

@@

```{r model1RunFreq}
# preProcess with near-zero value and kNN imputation
trnXFDpp <- preProcess(trnXFD, method = c('nzv', 'knnImpute'),
                       k = 3, knnSummary = median)
# Create processed predictors
trnXFDI <- predict(trnXFDpp, newdata = trnXFD)
trC <- trainControl(method = 'none',            # Using StepAIC
                    classProbs = TRUE,          # Classification
                    summaryFunction = prSummary # Precision/Recall
)
set.seed(487)
# Train Model
m1FreqFit <- train(x = trnXFDI, y = trnFreq, trControl = trC,
                   method = 'glmStepAIC', family = binomial(link = 'logit'),
                   direction = 'both', trace = 0)
# Print results
kable(summary(m1FreqFit$finalModel)$coefficients, digits = 3L,
      caption = "Model 1 Frequency Regression Output")
```

@@

```{r model1DummySev}
# Extract those observations with values > 0
trnSev <- trnX[TARGET_AMT > 0, TARGET_AMT]
# Set up dummy variables
trnDumSev <- dummyVars(TARGET_AMT ~ . - INDEX - TARGET_FLAG,
                        fullRank = TRUE, data = trnX[TARGET_AMT > 0])
# Create matrix of dummy values
trnXSD <- predict(trnDumSev, newdata = trnX[TARGET_AMT > 0])
# preProcess with near-zero value and kNN imputation
trnXSDpp <- preProcess(trnXSD, method = c('nzv', 'knnImpute'),
                       k = 3, knnSummary = median)
# Create processed predictors
trnXSDI <- predict(trnXSDpp, newdata = trnXSD)
```
```{r runModel1Sev}
trC <- trainControl(method = 'none', summaryFunction = defaultSummary,
                    classProbs = FALSE)
set.seed(878)
m1SevFit <- train(x = trnXSDI, y = trnSev, trControl = trC,
                  method = 'glmStepAIC', family = gaussian, direction = 'both',
                  trace = 0)
kable(summary(m1SevFit)$coefficients, digits = 3L,
      caption = "Model 1 Severity Regression Output")
```

@@

## Model Set #2
### Frequency Model

The frequency model is built to determine whether a vehicle is likely to crash. In Model 
Set #2, to work with complete separation in the logistic regression model, a Bayesian analysis 
is fitted. The Bayesian statistical model returns samples of the parameters of interest (the
"posterior" distribution) based on some "prior" distribution which is then updated by the 
data. Here, Cauchy distribution is accepted as a prior for parameters of the generalized 
linear model.

```{r model2RunFreq}
# Train Model Set #2
set.seed(525)
m2FreqFit <- train(x = m2.train.p, y = m2.train.r, 
                   trControl = trainControl(method = 'repeatedcv',
                                            classProbs = TRUE, 
                                            number = 10),
                   method = 'bayesglm', family = 'binomial')
# Results
kable(summary(m2FreqFit$finalModel)$coefficients, digits = 3L,
      caption = "Model 2 Frequency Regression Output")
```

#### Coefficient Discussion

Contributing coefficients are those which lie within the 95% level of significance. A 
positive coefficient indicates that as the value of the predictor increases, the response 
variable also tends to increase. A negative coefficient suggests that as the predictor 
increases, the response variable tends to decrease.

@@

Similar to Model Set #1, the multivariate linear regression model used the data set’s 
`TARGET_AMT` attribute as the response variable and the data was split to include only those
records that have a claim. The elastic net is a regularized regression method that linearly 
combines the L1 and L2 regularization penalties. Predictors, specifically qualitative variables,
were automatically transformed into dummy variables, which is important because `glmnet()` can 
only take numerical, quantitative inputs. Cross-validation is done on different combinations 
of $\lambda_1$ and $\lambda_2$ to find the best values. The hybrid elastic net regression is 
especially good at dealing with situations when there are correlations between parameters.

```{r model2RunSev}
# Train Model Set #2
temp <- split(m2.train.X, m2.train.X$TARGET_AMT > 0)[[2]] 
m2.train.r <- temp$TARGET_AMT
m2.train.p <- model.matrix(temp$TARGET_AMT~., temp[, -c(1:3)])[,-1]
set.seed(525)
m2SevFit <- train(x = m2.train.p, y = m2.train.r, 
                   trControl = trainControl(method = 'repeatedcv',
                                            number = 10),
                   method = 'glmnet',
                   tuneLength = 10,
                   preProcess = c("center", "scale"))
# Results
coeffs = coef(m2SevFit$finalModel, m2SevFit$bestTune$lambda)
kable(data.frame(cbind(variables = coeffs@Dimnames[[1]][coeffs@i+1], 
                       coef = round(coeffs@x,3))), 
      caption = "Model 2 Severity Regression Output")
```

#### Coefficient Disucssion

Once again, contributing coefficients are those which lie within the 95% level of significance. 
In this case, the linear model reduced the number of predictors that can predict the cost given 
an accident. The model fits extremely poor and does a mediocre job at explaining the variability 
within the training data. If interpretations are still to be made, first, the intercept itself 
suggests that with no information on the policyholder, given a vehicle accident, the likely claim 
value to the individual is $5,728. The other predictors suggest that the higher the bluebook value 
of the policyholder’s vehicle, the larger the insurance payout will be in the event of an accident. 
If they use the vehicle for commercial, they will also have a larger claim amount. While being a 
female resulted as a significant variable, its coefficient is near zero and will be ignored.

@@

```{r setUpCompTables}
freqCompTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
                            ACC = double(3),
                            F1 = double(3),
                            AUC = double(3))
sevCompTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
                           RMSE = double(3),
                           MAE = double(3),
                           adjR2 = double(3))
```

>>>>

```{r loadData, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(scales)
library(knitr)
library(caret)
library(RANN)
library(pROC)
library(mice)
library(arm)
library(corrplot)
library(data.table)

@@

```{r corrgram, fig.width=6, fig.height=4}
corrplot::corrplot(cor(DT[, ..nmtrnNUM], use = 'complete.obs'),
         method = 'ellipse', type = 'lower', order = 'hclust',
         hclust.method = 'ward.D2')
```

@@

```{r trainTestSplit}
# Create training and testing split
set.seed(642)
trnIDX <- createDataPartition(DT$TARGET_FLAG, p = 0.7)
trnX <- DT[trnIDX$Resample1, ]
tstX <- DT[!trnIDX$Resample1, ]
```

@@

```{r model1addVars}
# Model Set 1: Add OWN variable
trnX[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0),
                     levels = c(0, 1), labels = c('No', 'Yes'))]
tstX[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0),
                     levels = c(0, 1), labels = c('No', 'Yes'))]
```

## Model Set #2  
### Missing Data  
Like Model Set #1, the data preparation efforts for Model Set #2 also includes
the imputation of missing values for the predictor variables. However, because
it combines the standard linear regression and the nearest-neighbor imputation
approach, Model Set #2 elected to fill in missing values of a continuous
variable by using the predictive mean matching imputation method. Moreover,
there will be the derivation of three new binary factor variables.

@@

### New Variables  
For Model Set #2, new factor variables ($VARIABLE`.f`) were create as a way to 
better capture policyholder's profile in the binary logistic regression modeling
process.

@@

```{r m2clean}
# Model 2: Clean Training set
m2.train.X <- trnX
m2.train.X$HOME_VAL[is.na(m2.train.X$HOME_VAL)] <- 0
m2.train.X$OWN[is.na(m2.train.X$OWN)] <- "No"
m2.train.X$CAR_AGE = abs(m2.train.X$CAR_AGE) 
m2.train.r <- factor(m2.train.X$TARGET_FLAG, levels = c(0, 1), 
                     labels = c("NoClaim", "Claim"))
m2.train.X.impute <- mice(m2.train.X, method = 'pmm', print = FALSE)
m2.train.X <- complete(m2.train.X.impute)
m2.train.X$CAR_AGE.f <- factor(ifelse(m2.train.X$CAR_AGE > 1, 1, 0),
                              levels = c(0, 1), 
                              labels = c('1 Year or less', 'More than a Year'))
m2.train.X$HOMEKIDS.f <- factor(ifelse(m2.train.X$HOMEKIDS >= 1, 1, 0),
                              levels = c(0, 1), 
                              labels = c('None', 'At least 1'))
m2.train.X$KIDSDRIV.f <- factor(ifelse(m2.train.X$KIDSDRIV >= 1, 1, 0),
                              levels = c(0, 1), 
                              labels = c('None', 'At least 1'))
m2.train.p <- m2.train.X[, -c(1:3)]

@@

```{r model1DummyFreq}
# Model Set 1: Cast target as factor for caret purposes
trnFreq <- factor(trnX$TARGET_FLAG, levels = c(0, 1),
                  labels = c("NoClaim", "Claim"))
# Set up dummy variables
trnDumFreq <- dummyVars(TARGET_FLAG ~ . +
                          OWN * HOME_VAL * INCOME +
                          CLM_FREQ * OLDCLAIM * MVR_PTS +
                          CAR_USE * URBANICITY -
                         INDEX - TARGET_AMT,
                        fullRank = TRUE, data = trnX)
# Create matrix of dummy values
trnXFD <- predict(trnDumFreq, newdata = trnX)
```

@@

```{r model1RunFreq}
# Model Set 1: preProcess with near-zero value and kNN imputation
trnXFDpp <- preProcess(trnXFD, method = c('nzv', 'knnImpute'),
                       k = 3, knnSummary = median)
# Create processed predictors
trnXFDI <- predict(trnXFDpp, newdata = trnXFD)
trC <- trainControl(method = 'none',            # Using StepAIC
                    classProbs = TRUE,          # Classification
                    summaryFunction = prSummary # Precision/Recall
)
set.seed(487)
# Model Set 1: Train Model
m1FreqFit <- train(x = trnXFDI, y = trnFreq, trControl = trC,
                   method = 'glmStepAIC', family = binomial(link = 'logit'),
                   direction = 'both', trace = 0)
# Print results
kable(summary(m1FreqFit$finalModel)$coefficients, digits = 3L,
      caption = "Model 1 Frequency Regression Output")
```

@@

```{r model1DummySev}
# Model Set 1: Extract those observations with values > 0
trnSev <- trnX[TARGET_AMT > 0, TARGET_AMT]
# Set up dummy variables
trnDumSev <- dummyVars(TARGET_AMT ~ . - INDEX - TARGET_FLAG,
                        fullRank = TRUE, data = trnX[TARGET_AMT > 0])
# Create matrix of dummy values
trnXSD <- predict(trnDumSev, newdata = trnX[TARGET_AMT > 0])
# preProcess with near-zero value and kNN imputation
trnXSDpp <- preProcess(trnXSD, method = c('nzv', 'knnImpute'),
                       k = 3, knnSummary = median)
# Create processed predictors
trnXSDI <- predict(trnXSDpp, newdata = trnXSD)
```
```{r runModel1Sev}
# Model Set 1: Train Model
trC <- trainControl(method = 'none', summaryFunction = defaultSummary,
                    classProbs = FALSE)
set.seed(878)
m1SevFit <- train(x = trnXSDI, y = trnSev, trControl = trC,
                  method = 'glmStepAIC', family = gaussian, direction = 'both',
                  trace = 0)
kable(summary(m1SevFit)$coefficients, digits = 3L,
      caption = "Model 1 Severity Regression Output")
```

@@

## Model Set #2
### Frequency Model
The frequency model is built to determine whether a vehicle is likely to crash.
In Model Set #2, to work with complete separation in the logistic regression
model, a Bayesian analysis is fitted. The Bayesian statistical model returns
samples of the parameters of interest (the "posterior" distribution) based on
some "prior" distribution which is then updated by the data. Here, Cauchy
distribution is accepted as a prior for parameters of the generalized linear
model.

```{r model2RunFreq}
# Model Set 2: Train Frequency Model
set.seed(525)
m2FreqFit <- train(x = m2.train.p, y = m2.train.r, 
                   trControl = trainControl(method = 'repeatedcv',
                                            classProbs = TRUE, 
                                            number = 10),
                   method = 'bayesglm', family = 'binomial')
# Results
kable(summary(m2FreqFit$finalModel)$coefficients, digits = 3L,
      caption = "Model 2 Frequency Regression Output")
```

#### Coefficient Discussion 
Contributing coefficients are those which lie within the 95% level of
significance. A positive coefficient indicates that as the value of the
predictor increases, the response variable also tends to increase. A negative
coefficient suggests that as the predictor increases, the response variable
tends to decrease.

@@

### Severity Model  
Similar to Model Set #1, the multivariate linear regression model used the data
set’s `TARGET_AMT` attribute as the response variable and the data was split to
include only those records that have a claim. The elastic net is a regularized
regression method that linearly combines the L1 and L2 regularization penalties.
Predictors, specifically qualitative variables, were automatically transformed
into dummy variables, which is important because `glmnet()` can only take
numerical, quantitative inputs. Cross-validation is done on different
combinations of $\lambda_1$ and $\lambda_2$ to find the best values. The hybrid
elastic net regression is especially good at dealing with situations when there
are correlations between parameters.

```{r model2RunSev}
# Model Set 2: Train Severity Model
temp <- split(m2.train.X, m2.train.X$TARGET_AMT > 0)[[2]] 
m2.train.r <- temp$TARGET_AMT
m2.train.p <- model.matrix(temp$TARGET_AMT~., temp[, -c(1:3)])[,-1]
set.seed(525)
m2SevFit <- train(x = m2.train.p, y = m2.train.r, 
                   trControl = trainControl(method = 'repeatedcv',
                                            number = 10),
                   method = 'glmnet',
                   tuneLength = 10,
                   preProcess = c("center", "scale"))
# Results
coeffs = coef(m2SevFit$finalModel, m2SevFit$bestTune$lambda)
kable(data.frame(cbind(variables = coeffs@Dimnames[[1]][coeffs@i+1], 
                       coef = round(coeffs@x,3))), 
      caption = "Model 2 Severity Regression Output")
```

#### Coefficient Disucssion
Once again, contributing coefficients are those which lie within the 95% level
of significance. In this case, the linear model reduced the number of predictors
that can predict the cost given an accident. The model fits extremely poor and
does a mediocre job at explaining the variability within the training data. If
interpretations are still to be made, first, the intercept itself suggests that
with no information on the policyholder, given a vehicle accident, the likely
claim value to the individual is $5,728. The other predictors suggest that the
higher the bluebook value of the policyholder’s vehicle, the larger the
insurance payout will be in the event of an accident. If they use the vehicle
for commercial, they will also have a larger claim amount. While being a female
resulted as a significant variable, its coefficient is near zero and will be
ignored.

@@

```{r setUpCompTables}
# Create comparison tables
freqCompTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
                            ACC = double(3),
                            F1 = double(3),
                            AUC = double(3))
sevCompTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
                           RMSE = double(3),
                           MAE = double(3),
                           adjR2 = double(3))
```
